Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-07-05 patchy-selection-paper.tex

% ADD SOMETHING ABOUT:
% In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
% is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
% which depends fairly strongly on the selective disadvantage $s_m$ between patches.
% Equivalently, the rates are roughly equal if
% $R/\sigma = \log(1/(s_m A \mu))/\sqrt{2s_m}$,
% which gives the critical gap size past which adaptation will be mostly parallel
% in terms of selection coefficients, patch size, and mutation rate.

%commented out as wasn't quite sure what purpose this paragraph served.
%% the point was that most other work deals with bulk behavior, differential equations,
%% not tracking individuals as you have to do here.
% In this work, we have tried to emphasize an intuitive, individual-centered understanding of the process,
% determining, for instance, 
% the speed at which successful migrants traverse regions where they are maladapted,
% and the effective rate at which new families of migrant alleles appear
% at long distances from a patch.
% These quantities may be as useful for applied work
% as more abstract ones such as the probability that adaptation occurs by migration versus new mutation.

@@

\begin{figure}[ht!]
    \begin{center}
        % made in patchy-selection-plots.R
        \includegraphics{prob-establishment}
    \end{center}
    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
    in an abrupt transition (left) and a gradual transition (right).
    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
    and the selection coefficient interpolates linearly between these in the central hatched region.
    The number of offspring is Poisson.
    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
    \label{fig:prob_estab_calcs}
    }
\end{figure}

>>>>

Perhaps the most useful rule-of-thumb quantities we found were the following.
The effective rate of migration into an as-yet-unadapted patch from an already-adapted patch distance $R$ away 
-- the analogue of the mutational influx $\mu \rho A$ --
is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$.
Equivalently, the critical gap size between patches past which adaptation is likely independent
is $R = (\sigma/\sqrt{2s_m}) \log(1/(s_m A \mu))$.
Finally, successfully transiting migrant lineages move at rate $\sigma \sqrt{2s_m}$,
and so shared haplotype lengths between patches will be of order $R/\sigma \sqrt{2s_m}$.

@@

\begin{figure}[ht!]
    \begin{center}
        % made in patchy-selection-plots.R
        \includegraphics{prob-establishment}
    \end{center}
    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
    in an abrupt transition (left) and a gradual transition (right).
    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
    and the selection coefficient interpolates linearly between these in the central hatched region.
    The number of offspring is Poisson.
    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
    Note that the approximation used in the text is here $p_e \approx 2 s_p / \xi^2 = 0.1$,
    agreeing with the asymptotic value on the right.
    \label{fig:prob_estab_calcs}
    }
\end{figure}
